---
title: "coffee-rating"
author: "Richard Elek"
date: "2023-08-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidytuesdayR)
theme_set(theme_minimal())

tuesdata <- tidytuesdayR::tt_load('2020-07-07')
coffee_ratings <- tuesdata$coffee_ratings
```

## Coffee Ratings Analysis

In the last two years, coffee has became a hobby and I really enjoy being able to taste the different flavor profiles of different beans while adjusting various variables in order to make the perfect cup of coffee. Today, we'll be diving into one of those variables and looking at the different ratings of beans from around the world from the weekly Tidy Tuesday data project. <https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-07/readme.md>.


```{r Initial view of data, eval = FALSE}
glimpse(coffee_ratings)
```

```{r cleaning data}
coffee_ratings_cd <- coffee_ratings %>% 
  extract(grading_date, 'grading_year', '(\\d\\d\\d\\d)', convert = TRUE, remove = FALSE) %>% 
  filter(country_of_origin != 'NA') %>% 
  filter(altitude_mean_meters > 0)
```


### Top 20 Countries by Total Cupping Points
```{r Top Countries by Total Cupping Points}
top_countries <- coffee_ratings_cd %>%
  mutate(
    country_of_origin  = country_of_origin %>%
      fct_na_value_to_level() %>%
      fct_lump(n = 20) %>%
      fct_reorder(total_cup_points)) %>% 
  filter(total_cup_points > 0)

  ggplot(top_countries, aes(y = country_of_origin, 
                            x = total_cup_points)) +
  geom_boxplot(color = '#6F4E37') +
  labs(y = NULL,
       x = 'Total Cup Points',
       title = 'Top 20 Countries by Total Cupping Points')
```

Total cup points: the sum of the points given in each category: aroma, flavor, aftertaste, acidity, body, balance, uniformity, clean cup and, sweetness.

The box plot above shows the distribution of the top 20 countries by total cup points in descending order, including both Arabica and Robusta beans. It is no surprise to find regions in Africa and South America to be highly rated and at the top. United States has the highest total cup points, however it's important to note that it is skewed and greater distribution of points.


### Coffee Quality Between 2010 to 2018 
For the next part of the analysis, we'll be looking at the annual top scoring coffees. In order to get the year, the extract function was used to pull the year from the 'grading_date' column and then converted into an integer. This step was added to the cleaning stage. 
```{r Global Mean Points Over Time}
  coffee_ratings_cd %>%
  # Used to extract year from grading_year. Remove set to false keeps extracted column
  extract(grading_date, 'grading_year', '(\\d\\d\\d\\d)', convert = TRUE, remove = FALSE) %>% 
  group_by(grading_year) %>%
  filter(grading_year %in% c(2010:2018)) %>%
  summarise(points = mean(total_cup_points)) %>% 

  ggplot(aes(x = grading_year, 
             y = points)) +
  geom_line(colour = '#6f4e37') +
  geom_point() +
  labs(x = 'Years', 
       y = 'Total Cupping Points (mean)', 
       title = 'Global Mean Cuping Scores between 2010 to 2018')
```

Looking at the chart above, we could see a drop in mean score. Doing some additional research, I discovered 2012 was 'A year of extreme weather'. This included: droughts and flooding in regions of Brazil and Africa, and wildfires in Chile. Additional extreme weather events also took place in 2017. Therefore, this could be a contributing factor to the quality of coffee beans globally.


### Altitude and Quality of Coffee
```{r Altitude and Quality}
altitude_quality <- coffee_ratings_cd %>% 
  filter(altitude_mean_meters < 5000)

  ggplot(altitude_quality, aes(x = altitude_mean_meters, 
                               y = total_cup_points)) +
  geom_point(colour = '#6f4e37') +
  geom_smooth(method = 'lm') +
  labs(x = 'Mean Altitude (m)', 
       y = 'Total Cup Points', 
       title = 'Altitude and Quality')
```


```{r Correlation: Altitude and Quality Test}
cor.test(altitude_quality$total_cup_points, altitude_quality$altitude_mean_meters)
```
To confirm the relationship of the two variables being altitude and quality of coffee, a correlation test was ran. From this test, we could see a very small p-value of 0.0003 and the correlation value of 0.107889 suggesting that there is a statistical significant relationship, but weak positive correlation between the total cupping points and altitude. Other factors (weather, soil, etc.) play a bigger role in quality. 