---
title: "coffee-rating"
author: "Richard Elek"
date: "2023-08-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidytuesdayR)
theme_set(theme_minimal())

tuesdata <- tidytuesdayR::tt_load('2020-07-07')
coffee_ratings <- tuesdata$coffee_ratings
```

## Coffee Ratings Analysis

In the last two years, coffee has became a hobby and I really enjoy being able to taste the different flavor profiles of different beans while adjusting various variables in order to make the perfect cup of coffee. Today, we'll be diving into one of those variables and looking at the different ratings of beans from around the world from the weekly Tidy Tuesday data project. <https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-07/readme.md>.


```{r Initial view of data, eval = FALSE}
glimpse(coffee_ratings)
```

```{r cleaning data}
coffee_ratings_cd <- coffee_ratings %>% 
  extract(grading_date, 'grading_year', '(\\d\\d\\d\\d)', convert = TRUE, remove = FALSE) %>% 
  filter(country_of_origin != 'NA') %>% 
  filter(altitude_mean_meters > 0)
```

